Visualization and analysis of probability distributions of large temporal data

### 2022-02-02 Rmd/00-c-declaration.Rmd
date changed to 19th November
updated changes working on comments

---
output:
  pdf_document: default
  html_document: default
---
`r if (knitr::is_html_output()) '<!--'`

\newpage

# Thesis including published works declaration {-#ch-declare}

I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

Three publications are included in this thesis, one of which has been accepted for publication in the _Journal of Computational and Graphical Statistics_ (@Gupta2021-gravitas). Two others have not yet been submitted. "Exploring probability distributions across cyclic temporal granularities" is the central theme of the thesis. All of the papers in the thesis were conceptualized, developed, and written by myself, the student, while working in the Department of Econometrics and Business Statistics under the supervision of Professor Rob J Hyndman and Professor Dianne Cook. When co-authors are listed, it indicates that the work was produced through active collaboration between researchers and that their contributions to team-based research have been recognized. The following table details the work, including my and my fellow co-authors' contributions.

\begin{table}
\centering\footnotesize\tabcolsep=0.12cm
\begin{tabular}{|p{1cm}|p{2cm}|p{1.5cm}|p{3.5cm}|p{3.5cm}|p{1.5cm}|}
\hline
\RaggedRight\textbf{Thesis Chapter}  & \RaggedRight\textbf{Publication Title}  & \RaggedRight\textbf{Status (published, in press, accepted or returned for revision)}  & \RaggedRight\textbf{Nature and} {\%} \RaggedRight\textbf{of student contribution} & \RaggedRight\textbf{Co-author name(s) Nature and} {\%} \RaggedRight\textbf{of Co-author’s contribution} & \RaggedRight\textbf{Co-author(s), Monash student Y/N} \\ \hline
2 & Visualizing probability distributions across bivariate cyclic temporal granularities & Published & 70\%. Concept, methodology development, writing first draft and software development & (1) Rob J Hyndman, Concept, methodology and software development, writing 12\% (2) Dianne Cook, Concept, methodology and software development, writing 12\% (3) Antony Unwin, Concept and software development 6\% & \multicolumn{1}{c|}{N} \\
\hline
\end{tabular}
\end{table}

I have renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.

@@

<!-- eval = file.exists("img/signature.png") -->

\textbf{Date:} `r Sys.Date()`

`r if (knitr::is_html_output()) '-->'`

\vspace{2em}
\noindent
The undersigned hereby certify that the above declaration correctly reflects the nature and extent of the student's co-authors' contribution to this work.
In instances where I am not the responsible author I have consulted with the responsible author to agree on the respective contributions of the authors.

\textbf{Main Supervisor name:} Rob J Hyndman

\textbf{Main Supervisor signature:}

\textbf{Date:} `r Sys.Date()`

>>>>

# Declaration {-#ch-declare}

This thesis is an original work of my research and contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

\vspace{4em}

@@

\textbf{Date:} `r as.Date("2021-11-19")`

### 2022-02-02 Rmd/00-e-preface.Rmd
updated changes working on comments

Chapter \@ref(ch-gravitas) has been accepted by the _Journal of Computational and Graphical Statistics_. It has won the ACEMS Business Analytics Prize in 2020. The accompanying R package `gravitas` is on CRAN. Chapter \@ref(ch-hakear) and Chapter \@ref(ch-gracsr) are yet to be submitted.

>>>>

Chapter \@ref(ch-gravitas) has been published online at the [_Journal of Computational and Graphical Statistics_](https://www.tandfonline.com/eprint/NXKCSZHIBFXCJD2NKB68/full?target=10.1080/10618600.2021.1938588). It has won the ACEMS Business Analytics Prize in 2020. The accompanying R package `gravitas` is on CRAN. Chapter \@ref(ch-hakear) and Chapter \@ref(ch-gracsr) are yet to be submitted.

### 2022-02-02 Rmd/02-gravitas.Rmd
updated changes working on comments

\noindent In the Mayan calendar (Table \@ref{tab:tab-mayan}), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.

>>>>

\noindent In the Mayan calendar (\autoref{tab:tab-mayan}), kin-of-tun or kin-of-baktun are examples of multiple-order-up granularities and single-order-up granularities are kin-of-uinal, uinal-of-tun etc.

### 2022-02-02 Rmd/04-gracsr.Rmd
updated changes working on comments

<!-- time series clustering and its challenges -->
Time series clustering is the process of unsupervised partitioning of $n$ time series data into $k$ ($k<n$) meaningful groups such that homogeneous time series are grouped together based on a certain similarity measure. The time series features, length of time series, representation technique, and, of course, the purpose of clustering time series all influence the suitable similarity measure or distance metric to a meaningful level. The three primary methods to time series clustering [@liao2005clustering] are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based). The efficacy of distance-based techniques is highly dependent on the distance measure utilized. Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data. Commonly used distance-based similarity measures as suggested by a decade review of time series clustering approaches [@Aghabozorgi2015-ct] are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances [@liao2007clustering] but on vector time series data. Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features. DTW works well with time series of different lengths [@corradini2001dynamic], but it is incapable of handling missing observations. Surprisingly, probability distributions, which may reflect the inherent temporal structure of a time series, have not been considered in determining time series similarity.

>>>>

<!-- time series clustering and its challenges -->
Time series clustering is the process of unsupervised partitioning of $n$ time series data into $k$ ($k<n$) meaningful groups such that homogeneous time series are grouped together based on a certain similarity measure. The time series features, length of time series, representation technique, and, of course, the purpose of clustering time series all influence the suitable similarity measure or distance metric to a meaningful level. The three primary methods of time series clustering [@liao2005clustering] are algorithms that operate directly with distances or raw data points in the time or frequency domain (distance-based), with features derived from raw data (feature-based), or indirectly with models constructed from raw data (model-based). The efficacy of distance-based techniques is highly dependent on the distance measure utilized. Defining an appropriate distance measure for the raw time series may be a difficult task since it must take into account noise, variable lengths of time series, asynchronous time series, different scales, and missing data. Commonly used distance-based similarity measures as suggested by a review of time series clustering approaches [@Aghabozorgi2015-ct] are Euclidean, Pearson's correlation coefficient and related distances, Dynamic Time Warping (DTW), Autocorrelation, Short time series distance, Piecewise regularization, cross-correlation between time series, or a symmetric version of the Kullback–Liebler distances [@liao2007clustering] but on vector time series data. Among these alternatives, Euclidean distances have high performance but need the same length of data over the same period, resulting in information loss regardless of whether it is on raw data or a smaller collection of features. DTW works well with time series of different lengths [@corradini2001dynamic], but it is incapable of handling missing observations. Surprisingly, probability distributions, which may reflect the inherent temporal structure of a time series, have not been considered in determining time series similarity.
